---
title: "Coursera_Practical ML"
author: "durvesh"
date: "5/15/2021"
output:
  html_document:
    df_print: paged
---

#Assignment 

#Using devices such as Jawbone Up, Nike FuelBand, and Fitbit, it is now possible to collect a large amount of data about personal activity relatively inexpensively. The aim of this project is to predict the manner in which participants perform a barbell lift. The data comes from http://groupware.les.inf.puc-rio.br/har wherein 6 participants were asked to perform the same set of exercises correctly and incorrectly with accelerometers placed on the belt, forearm, arm, and dumbell.

#For the purpose of this project, the following steps would be followed:

#1.Import the Data 
#2.Removing Columns that are not relevant
#3.Using Random Forest 
#4.Predicting Test Set Output for 20 cases 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(tidyverse)
library(future)
library(randomForest)

```


```{r}

## Loading the training and the test data from the given url 

train<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
test<-read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))


```


```{r}
# We check the size of the training and test sets 

object.size(train)
object.size(test)


```

```{r}
# For the classification problem we find number of classes to predict 
unique(train$classe)
```

```{r}
train<-train[,-c(1:6)] # removing the id columns that are not relevant train
test<-test[,-c(1:6)]   # removing the id columns that are not relevant test

var<-nearZeroVar(train) # find the variables with zero variance 
var<-var[-26] # retaining the classe variable 
t1<-train %>% select(-all_of(var))
t2<-select_if(t1,function(x) sum(is.na(x)) <=1) #dropping columns that contain nas 
```


```{r}
plan(multisession)  # parallel processing to speed up
t2$classe <-as.factor(t2$classe) # converting the classification var to factor
model=randomForest(classe~., data=t2, method='class') 


```

```{r}
model # printing the results of the model

```

```{r}
vec_names<-names(t2)
vec_names<-vec_names[-length(vec_names)] # applying the same subset from train to test set columns  removing the classe columns 
test_1<-test %>% select(all_of(vec_names))


```



```{r}
base_predict<-predict(model,test_1) # creating a prediction based on model
```

```{r}
base_predict # printing out the results of prediction 

```

